---
title: "STAT 421 Lab 5"
author: "Your Name"
date: "November 1, 2019"
output: html_document
---

This lab/qz deals with the GLSD and its properties. The following data comes from problem 4.35. You may copy/paste the following data and code from the file lab5_supp.txt on the course website, but study it carefully to assure you understand what it does:

RUN THIS CODE TO GET THE DATA
```{r}
rm(list=ls(all=TRUE))
p = 5
y.m = matrix(nrow=p,ncol=p)
y.m[1,]=c(26,16,19,16,13)
y.m[2,]=c(18,21,18,11,21)
y.m[3,]=c(20,12,16,25,13)
y.m[4,]=c(15,15,22,14,17)
y.m[5,]=c(10,24,17,17,14)
y = as.vector(t(y.m))

A = as.factor(c(rep(1:p,each=p))) 	# row-factor (Order of Assembly).
D = as.factor(rep(c(1:p),p)) 	# col-factor (Operator). Books convention (p.166)
B = as.factor(c(
              1,2,3,4,5,   		
              2,3,4,5,1,
              3,4,5,1,2,
              4,5,1,2,3,
              5,1,2,3,4)) # Greek factor (treatment)

C = as.factor(c(
              1,2,3,4,5,  		
              3,4,5,1,2,
              5,1,2,3,4,
              2,3,4,5,1,
              4,5,1,2,3)) # Latin factor

cbind(A,B,C,D,y)		 # Always, visually confirm data are correctly read in.
```

RUN THIS CODE
```{r}
lm.1 = lm(y~A+B+C+D)
summary.aov(lm.1)
```

Just for fun, run the last line again, with the order of B, C changed.
```{r}
lm.2 = lm(y~A+C+B+D)
summary.aov(lm.2)
```

*a) Write code to compute SSE By R. Call it SSE_0.*
```{r}
# write your code here
summary1 <- summary.aov(lm.1)[[1]]
SSE_0 <- summary1$`Sum Sq`[5]
SSE_0
```

*b) Write code to compute SSA, SSB, SSC, SSD, (NOT SSE), all BY HAND. Report them at the end of your code with c(SSA, SSB, SSC, SSD, SSE).* 
Hint: You'll have to find the various y-sums, but then you can use the following formulas
  SST = sum(y^2) - y....^2/p^2,
  SSA = sum(yi...^2)/p - y....^2/p^2
  etc.
```{r}
# write your code here
y.... <- sum(y)
yi... <- c(sum(y[A==1]), sum(y[A==2]), sum(y[A==3]), sum(y[A==4]), sum(y[A==5]))
y.j.. <- c(sum(y[B==1]), sum(y[B==2]), sum(y[B==3]), sum(y[B==4]), sum(y[B==5]))
y..k. <- c(sum(y[C==1]), sum(y[C==2]), sum(y[C==3]), sum(y[C==4]), sum(y[C==5]))
y...l <- c(sum(y[D==1]), sum(y[D==2]), sum(y[D==3]), sum(y[D==4]), sum(y[D==5])) 

SST <- sum(y^2) - y....^2/p^2

SSA <- sum(yi...^2)/p - y....^2/p^2
SSB <- sum(y.j..^2)/p - y....^2 / p^2
SSC <- sum(y..k.^2)/p - y....^2 / p^2
SSD <- sum(y...l^2)/p - y....^2 / p^2
c(SSA, SSB, SSC, SSD)

SST - (SSA + SSB + SSC + SSD + SSE_0)

```
Confirm that SST - SSA - SSB - SSC - SSD - SSE_0 = 0, i.e., that the ANOVA decomposition holds.

*c)  Now, change the design of C as shown below, and produce the ANOVA table, By R.*
```{r}
C = as.factor(c(
              1,2,3,4,5,        # Latin factor
              3,4,5,1,2,
              5,1,2,3,4,
              4,5,1,2,3, 
              2,3,4,5,1))
```

```{r}
# write your code here
lm.3 <- lm(y~A+B+C+D)
summary.aov(lm.3)

```

It should not be surprising that you get different SS values compared to above; after all, it's a different design. But, there is another, much bigger problem!  To see what,

*d) Write code to compute SSE of this design, By R, and call it SSE_1.*
```{r}
# write your code here
summary1 <- summary.aov(lm.1)[[1]]
SSE_1 <- summary.aov(lm.3)[[1]]$`Sum Sq`[5]

```

*e) Use the "By hand" code you made in part b to report c(SST, SSA, SSB, SSC, SSD)*
```{r}
# write your code here
y.... <- sum(y)
yi... <- c(sum(y[A==1]), sum(y[A==2]), sum(y[A==3]), sum(y[A==4]), sum(y[A==5]))
y.j.. <- c(sum(y[B==1]), sum(y[B==2]), sum(y[B==3]), sum(y[B==4]), sum(y[B==5]))
y..k. <- c(sum(y[C==1]), sum(y[C==2]), sum(y[C==3]), sum(y[C==4]), sum(y[C==5]))
y...l <- c(sum(y[D==1]), sum(y[D==2]), sum(y[D==3]), sum(y[D==4]), sum(y[D==5])) 

SST <- sum(y^2) - y....^2/p^2

SSA <- sum(yi...^2)/p - y....^2/p^2
SSB <- sum(y.j..^2)/p - y....^2 / p^2
SSC <- sum(y..k.^2)/p - y....^2 / p^2
SSD <- sum(y...l^2)/p - y....^2 / p^2
c(SSA, SSB, SSC, SSD)

SST - (SSA +SSB + SSC + SSD + SSE_1)
```
Confirm that SST - SSA - SSB - SSC - SSD - SSE_1 is NOT zero.

*f) Why is it nonzero? Give an explanation based on everything we have been discussing in lectures and here.*

Via combinations, we can see there are identical pairs exist in the new GLSD, therefore, these two LSD's are not othorgonal to each other. i.e. the SScross's are non zero between i, k or j, l.

Because the cross terms are non-zero.

*g) Just for fun, find the ANOVA tables for this latest design, By R, but by switching the order of*
- A and B,
- C and D,
- A and D,
- B and C.
```{r}
# switch A and B
lm.4 <- lm(y~B+A+C+D)
summary.aov(lm.4)

```

```{r}
# switch C and D
lm.5 <- lm(y~A+B+D+C)
summary.aov(lm.5)

```

```{r}
# switch A and D
lm.6 <- lm(y~D+B+C+A)
summary.aov(lm.6)

```

```{r}
# switch B and C
lm.7 <- lm(y~A+C+B+D)
summary.aov(lm.7)

```

Later, check the solutions to find out why all this is happening.

When the two LSD's are orthorgonal (the original one), change the order of BC does not change sum of squares, because SScross's for each terms of that GLSD is zero.

Since in the latest model, SScross's between i, k or j, l are non-zero, change the order of B,C will change the values of SScross's, and finally affects the values of SSC and SSB. 

